Automated Lensing Learner - I: An automated Strong Lensing Identification Pipeline

### 2017-03-07 ms.tex
Additional figure captions added

\documentclass{emulateapj}
\usepackage{amsmath} 
\usepackage{apjfonts} 
\usepackage{amssymb} 
\usepackage{xspace}
\usepackage{hyperref} 
\usepackage{natbib} 
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{epsf}
\usepackage{subfigure}
\bibliographystyle{apj_ads}
\usepackage[usenames,dvipsnames]{xcolor}
\newcommand{\comment}[1]{{\color{red} #1}}
\newcommand{\change}[1]{{\it\color{purple} #1}}
\newcommand{\todo}[1]{{\bf\color{blue} #1}}
%%%%%%%%%% User defined symbol %%%%%%
\def\gsim{\gtrsim}
\def\lsim{\lesssim}    
\def\Msun{M_\odot}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%-------------------------------------------------%
\begin{abstract} 
Gravitational lensing offers a direct probe of the underlying mass
distribution of lensing systems, a window to the high redshift
universe, and a geometric probe of cosmological models.  The advent of
large scale surveys such as the Large Synoptic Sky Telescope and
Euclid has prompted a need for automatic and efficient identification
of strong lensing systems.  We present {\em (ALL) Automated Lensing
  Learner}, a strong lensing identification pipeline that will be
publicly released as open source software.  In this first application,
we employ a fast feature extraction method, Histogram of Oriented
Gradients (HOG), to capture edge patterns that are characteristic of
strong gravitational arcs in galaxy-galaxy lensing.  We use logistic
regression to train a supervised classifier model on the HOG of HST-
and LSST-like images.  We use an area under the curve (AUC) of a
Receiver Operating Characteristic curve to assess model performance;
$AUC=1.0$ is an ideal trained classifier, and $AUC=0.5$ does not
provide any classification information.  Our best performing models
exhibit an AUC of \change{XX} for an HST-like sample.  However, for
one orbit of LSST, our model only reaches an AUC of \change{XX}.  In
absence of the lens galaxy in one orbit of LSST-like data, the AUC
improves to \change{XX}.  For 10-year mock LSST observations, the
respective AUC with and without the lens galaxy improves to
\change{XX} and \change{XX}.  Our results show an efficient and
effective methods for automated strong lensing identification.  The
results also emphasize the need for further development and testing of
additional tools to maximize use of ground-based surveys.
\end{abstract}
%-------------------------------------------------%

@@

% Summary of HST and LSST images
We use mock {\em Hubble Space Telescope} (HST) and {\em Large Synoptic
  Sky Telescope} (LSST) images generated with PICS (Pipeline for
Images of Cosmological Strong lensing) \citep{li_etal16}.  These
images include cluster-member galaxies and foreground stars.  Half of
the images used are strong lensing systems.  Here, lensed galaxy
images are ray-traced images of actual galaxies extracted from deep
Hubble Space Telescope observations.  The final mock images are
``observed'' with a realistic point spread function corresponding to
modeled detector artifacts for bright stars.  See \cite{li_etal16} for
more details.

@@

To perform our analysis, we have used a variety of tools from {\em
  Scikit-learn} \citep{pedregosa_etal12}.  We outline the
identification pipeline in Figure~\ref{fig:pipeline}.  The first step
of our pipeline consists of a feature extraction stage, where our
feature vector is a {\em histogram of oriented gradients}
\citep{dalalandtriggs_05}.  We describe the method in
Section~\ref{sec:hog}.  We then use a machine learning algorithm to
train a classifier model on a subset of our images.  We have explored
both {\em Logistic Regression} (LogReg) and {\em Support Vector Machine}
(SVM) machine learning algorithms.  We test our trained model on an
independent subset of the images to assess the model performance.

Both the feature extrator, HOG, and the classifier contain a range of
parameters, which must be tested and optimized for peak model
performance.  We use {\em GridSearchCV} from {\em Scikit-learn} to
select appropriate parameters, and discuss these results in
Section~\ref{sec:gridsearch}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]\label{fig:pipeline}
\begin{center}
\includegraphics[scale=0.55]{figures/supervised_classification_workflow.png}
\caption{** Cartoon of pipeline ** Mention that median filtering generates
visual artifacts that confuse the model. http://www.nltk.org/book/ch06.html}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%-------------------------------------------------%
\subsection{Feature Extractor: Histogram Oriented Gradients}\label{sec:hog}
%-------------------------------------------------%
Histogram oriented gradients (HOG) is a feature extraction method that
computes centered horizontal and vertical gradients with no smoothing.
Originally created for human detection, .... This computes the
gradient orientation and magnitudes in an image by first dividing the
image into $n\times n$ blocks of X\% overlap.  The gradient
orientation is then quantized into m bins. ``(optional) global image
normalisation computing the gradient image in x and y computing
gradient histograms normalising across blocks flattening into a
feature vector''\citep{dalalandtriggs_05}.

% Need to look into edge orientation histograms, scale-invariant
% feature transform descriptors, shape contexts.

% NEED TO FEED THE DESCRIPTORS TO SOMETHING WITH SUPERVISED LEARNING.
% (e.g. SVM, Neural network).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure*}[t]\label{fig:mockimages}
\begin{center}
\includegraphics[width=2\columnwidth]{figures/tiles/compilations/hog_telescope_compilation.pdf}
\caption{Left to right show example image transforms of mock images
  from Figure~\ref{fig:mockimages} with a visualized histogram of
  oriented gradients.  The image transform picks up edge features,
  with arc features showing up as edges across radial orientations.}
\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%-------------------------------------------------%
\subsection{Machine Learning Algorithm: Logistic Regression}\label{sec:logreg}
%-------------------------------------------------%

Logisitic regression is a linear classifier, which finds the optimal
hyperplane between two classes of data points.  This hyperplane
defines the optimal model.  We must regularize the model to prevent
overfitting.

In regularization, the goal is to minimize a cost function with
respect to ....  Larger values of $C_\text{LogReg}$ correspond to
increasing model complexity.  Figure~\ref{fig:regularization} shows
the accuracy of a model for selected HOG parameters as a function of
the regularization parameter, $C_\text{LogReg}$.  We plot the
regularization dependency for HOG parameterization of
$N_\text{orient}=9$, $PPC=(16, 16)$, $CPB=(1,1)$.

The accuracy of a model can be defined as either the score of the
model on a test set, or the AUC for the ROC curve.

% Regularization coefficient:
% http://openclassroom.stanford.edu/MainFolder/DocumentPage.php?course=MachineLearning&doc=exercises/ex5/ex5.html

%-------------------------------------------------%
\subsection{Machine Learning Algorithm: Support Vector Machine}
%-------------------------------------------------%

%-------------------------------------------------%
%\subsection{Deep Learning Algorithm}
%-------------------------------------------------%

%-------------------------------------------------%
\section{Results}
\label{sec:results}
%-------------------------------------------------%
%-------------------------------------------------%
\subsection{Optimized pipeline parameters}\label{sec:gridsearch}
%-------------------------------------------------%

\begin{table}
\caption{Grid Search of Pipeline Parameters}
\begin{center}
\begin{tabular}{cccccc}
\hline \\ [-0.2cm]
N$_{orient}$ & Pixels/Cell & Cells/Block & C$_{reg}$ & HOG size & Score \\ [0.2cm]
\hline \\ [-0.2cm]
\multicolumn{6}{c}{(a) HST-like data} \\ [0.2cm]
\hline \\ [-0.2cm]
9 & (4, 4) & (3, 3) & 50. & &$0.76181\pm0.00491$\\ [0.2cm]
9 & (4, 4) & (1, 1) & 50. & &$0.83403\pm0.00547$\\ [0.2cm]
9 & (8, 8) & (3, 3) & 50. & &$0.85417\pm0.00851$\\ [0.2cm] 
9 & (8, 8) & (1, 1) & 50. & &$0.86389\pm0.00687$\\ [0.2cm] 
9 & (16, 16) & (1, 1) & 50. & & $0.88542\pm0.00680$ \\ [0.2cm]
9 & (16, 16) & (3, 3) & 10. & &$0.89306\pm0.00687$ \\ [0.2cm]
8 & (16, 16) & (3, 3) & 50. & &$0.90000\pm0.00170$ \\ [0.2cm]
9 & (16, 16) & (3, 3) & 50. & &$0.90139\pm0.00804$ \\ [0.2cm]
\hline \\ [-0.2cm]
\multicolumn{6}{c}{(b) LSST-like data} \\ [0.2cm]
\hline \\ [-0.2cm]
& & & & &\\
& & & & &\\ [0.2cm]
\hline
\end{tabular}
\end{center}
\label{tab:gridsearch}
\tablecomments{Panel (a) shows the results of a grid search across a
  range of HOG parameters and regression coefficient with training and
  test sets of size 1440 with the HST-like mock dataset.  Panel (b)
  shows the corresponding results for training and test sets of size
  5760 with the LSST mock dataset.  We explore different HOG
  parameters in each dataset due to resolution and image size
  differences.}
\end{table}

We run a grid search across parameters that should reasonably sample
the arc edges in either the HST- or LSST-like mock observations.
Recall, the HST-like images are $300\times300$ pixels per image, while
the LSST-like mock observations are $45\times45$ pixels per image.  In
general, must run a grid search in order to best optimize the pipeline
parameters.

We must first estimate the size of a cell that will contain a coherent
arc feature.  To first order approximation, subdivisions of cells that
are 1/100th the area of the entire image should contain coherent arc
edges that span an elongated shape within arc-containing cells.
Therefore, we perturb the pixels per cell parameter about (30, 30) for
the HST-like images, and (4, 4) for the LSST-like images.

Next, the cells per block parameter determines the normalization of
each cell with respect to the neighboring cell.  In general, this will
downweight arc-like edges in cells that neighbor very bright cells,
such as cells that cover the BCG.  We therefore vary the cells per
block parameter between (1, 1) and (3, 3).

The number of orientations will determine the sampling of rounded
edges.  For example, if we only have two orientations, an arc-like
feature in a cell will appear as .... However, contributions from a
cluster or line-of-sight galaxy in the same cell will ... the
normalizations in the histogram.  The resolution of the overall image
will also limit the additional information that an increase in
$N_\text{orient}$ will provide.  For the LSST-like data,
$N_\text{orient}=9$ compared with $N_\text{orient}=5$ only
results in a $???\%$ improvement.  In the case of the HST-like data,
$N_\text{orient}=9$ compared with $N_\text{orient}=5$
results in a $???\%$ improvement.

Note, we also include the size of the extracted features, e.g. the
length of the HOG feature vector.  A longer feature vector will
increase the amount of time required to train the model.
Additionally, for fixed memory restrictions, there is a tradeoff
between the length of the feature vector and the size of the training
set.  The logistic regression coefficient will also increase the
training time, \todo{since larger coefficients correspond to a more
  singular matrix}.  We discuss the score dependence on logistic
regression coefficient in Section\ref{sec:regularization}.

%-------------------------------------------------%
\subsection{Receiver Operating Characteristic}
%-------------------------------------------------%

@@

%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
%% \begin{figure}[t]\label{fig:ROCHST}
%% \begin{center}
%% \includegraphics[width=\columnwidth]{figures/ROC_HST_HOG.png}
%% \caption{Red, green, blue: ROC curves for a models trained on the
%%   HST-like data with the Logistic Regression algorithm.  The magenta
%%   solid (dotted) curve show the ROC curve for the LSST-like data with
%%   (without) the BCG.  The corresponding parameters are labeled: N for
%%   number of orientations, ppc for pixels-per-cell, cpb for
%%   cells-per-block, and C for best regularization parameter.  The AUC
%%   for each model is also labeled in the legend, where an AUC close to
%%   1 is a near-ideal model, and an AUC close to 0.5 is a model that
%%   performs as well as a coin-flip.}
%% \end{center}
%% \end{figure}
%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%            
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]\label{fig:ROCcompilation}
\begin{center}
\includegraphics[width=\columnwidth]{figures/metrics/ROC/ROC_compilation.pdf}
\caption{Red, blue, green: ROC curves for models trained on our whole
  10,000 training set and tested on our holdout set of 1,000.  These
  respectively correspond to the HST, LSST 1 year, and LSST 10 year
  data.  The solid lines are for data that include the lensing central
  galaxy, and the dashed lines for the data where there is no lensing
  galaxy, mimicking an ideal removal of the lens.  Model performance
  can be summarized by the area under the curve (AUC), labeled in the
  legend. $AUC=1$ is a perfect model, and $AUC=0.5$ is a useless
  model.}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% Logistic regression and regularization
As described in Section~\ref{sec:logreg}, LogReg trains a model with
complexity determined by the regularization parameter coefficient,
$C_\text{LogReg}$.  Increasing model complexity (larger
$C_\text{LogReg}$), will better describe features in the training
set[wc?].  However, an overly complex model will overfit the training
set at the expense of its performance on any independent test set.
The regularization parameter ultimately defines the model performance,
and we must perform a parameter search to identify the optimal value
for $C_\text{LogReg}$.

@@

%-------------------------------------------------%
\subsection{Effects of Dataset Size on Model Performance}\label{sec:datasize}
%-------------------------------------------------% 
%-------------------------------------------------% 
\subsubsection{Effects of Training Set Size}
%-------------------------------------------------% 
In this section, we show the effects of training set size on model
performance.  Figure~\ref{fig:trainsize} shows the AUC and the train
time depend on the log of the size of the training set.  In solid blue
lines, we see that the AUC improves almost linearly with the log of
the training size set.  In dotted red, we see that the train time
roughly scales logarithmically with the log of the train size, with an
exception of the data point corresponding to train size of 2000.

@@

%% \caption{LSST-like images with a BCG. These illustrate the same image
%%   classification paradigms as Figure~\ref{fig:HSTclassificationtile}.
%%   Note, the larger PSF of LSST reduces the resolution of images,
%%   making the arc-like morphology of faint images less significant and
%%   increases the appearance of image-BCG blending.}
%% \label{fig:LSSTBCGclassificationtile}
%% \end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%            
\subsection{BCG removed}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
%% \begin{figure*}[t]
%%   \centering

@@

%-------------------------------------------------%
\subsubsection{Performance dependence on lens-model parameters}
\label{sec:peformance_lensmodel}
%-------------------------------------------------%

>>>>

\documentclass{emulateapj}
\usepackage{amsmath} 
\usepackage{apjfonts} 
\usepackage{amssymb} 
\usepackage{xspace}
\usepackage{hyperref} 
\usepackage{natbib} 
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{epsf}
\usepackage{subfigure}
\bibliographystyle{apj_ads}
\usepackage[usenames,dvipsnames]{xcolor}
\newcommand{\comment}[1]{{\color{red} #1}}
\newcommand{\wording}[1]{{\it\color{purple} #1}}
\newcommand{\todo}[1]{{\bf\color{blue} #1}}
%%%%%%%%%% User defined symbol %%%%%%
\def\gsim{\gtrsim}
\def\lsim{\lesssim}    
\def\Msun{M_\odot}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%-------------------------------------------------%
\begin{abstract} 
Gravitational lensing offers a direct probe of the underlying mass
distribution of lensing systems, a window to the high redshift
universe, and a geometric probe of cosmological models.  The advent of
large scale surveys such as the Large Synoptic Sky Telescope and
Euclid has prompted a need for automatic and efficient identification
of strong lensing systems.  We present {\em (ALL) Automated Lensing
  Learner}, a strong lensing identification pipeline that will be
publicly released as open source software.  In this first application,
we employ a fast feature extraction method, Histogram of Oriented
Gradients (HOG), to capture edge patterns that are characteristic of
strong gravitational arcs in galaxy-galaxy lensing.  We use logistic
regression to train a supervised classifier model on the HOG of HST-
and LSST-like images.  We use an area under the curve (AUC) of a
Receiver Operating Characteristic curve to assess model performance;
$AUC=1.0$ is an ideal trained classifier, and $AUC=0.5$ does not
provide any classification information.  Our best performing models
exhibit an AUC of \todo{XX} for an HST-like sample.  However, for one
orbit of LSST, our model only reaches an AUC of \todo{XX}.  In absence
of the lens galaxy in one orbit of LSST-like data, the AUC improves to
\todo{XX}.  For 10-year mock LSST observations, the respective AUC
with and without the lens galaxy improves to \todo{XX} and \todo{XX}.
Our results show efficient and effective methods for automated strong
lensing identification with \wording{scalable linear methods that are
  easy to parallelize with existing open source tools.}
\end{abstract}
%-------------------------------------------------%

@@

% Summary of HST and LSST images
We use mock {\em Hubble Space Telescope} (HST) and {\em Large Synoptic
  Sky Telescope} (LSST) images generated with PICS (Pipeline for
Images of Cosmological Strong lensing) \citep{li_etal16} and
LensPop\citep{collett_etal15}.  These images include cluster-member
galaxies and foreground stars.  Half of the images used are strong
lensing systems.  Here, lensed galaxy images are ray-traced images of
actual galaxies extracted from deep Hubble Space Telescope
observations.  The final mock images are ``observed'' with a realistic
point spread function corresponding to modeled detector artifacts for
bright stars.  See \cite{li_etal16} for more details.

@@

To perform our analysis, we have used tools from {\em Scikit-learn}
\citep{pedregosa_etal12}.  We outline the identification pipeline in
Figure~\ref{fig:pipeline}, which is a general description of
supervised classification.  Supervised classification is a class of
machine learning where the labels of classification in the training
set are known.  In our case, the labels are lens and non-lens
containing images.

The first step of our pipeline consists of a feature extraction stage,
where our feature vector is a {\em histogram of oriented gradients}
\citep{dalalandtriggs_05} that quantifies edges in the image.  We
describe the method and parameter search in Section~\ref{sec:hog}.  We
then use {\em Logistic Regression} (LR), a machine learning algorithm
described in Section~\ref{sec:LR}, to train a classifier model on a
subset of our images.  LR requires a parameter search over the
regression coefficient, $C_{LogReg}$, which we explore in
Section~\ref{sec:regularization}.  We briefly comment that tests with
\todo{{\em Support Vector Machine} (SVM) as an alternative machine
  learning algorithm yielded negligible performance improvement, and
  increased computation cost.  We do not include SVM in our final
  analysis and comparisons.}

Both the feature extrator, HOG, and the linear classifier, LR, contain
parameters, which must be tested and optimized for peak model
performance.  We use {\em GridSearchCV} from {\em Scikit-learn} to
select cross-validated parameters, and discuss this step of our
methodology in Section~\ref{sec:gridsearch}.

The second step of our analysis is to test our trained model on an
independent subset of the images to assess the model performance.
Here, we evaluate the model on each test image, predicting a
likelihood (``score'') between 0 and 1 that this is a lens containing
system.  This ``holdout set'' is not used in any of our parameter
searches to keep our test metric independent of tuning.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]\label{fig:pipeline}
\begin{center}
\includegraphics[scale=0.55]{figures/supervised_classification_workflow.png}
\caption{** Cartoon of pipeline ** from
  http://www.nltk.org/book/ch06.html}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%-------------------------------------------------%
\subsection{Feature Extractor: Histogram Oriented Gradients}\label{sec:hog}
%-------------------------------------------------%
% HOG intro - 
Originally created for human detection in computer vision, histogram
oriented gradients (HOG) is a feature extraction method that computes
centered horizontal and vertical gradients. HOG is relatively robust
to noise in the image, and is a fairly fast transform that describes
edges.  Details can be found in \citet{dalalandtriggs_05}, but we
describe the procedure here.  The end result of HOG is a one
dimensional histogram computed as follows.

% Division of the image  
HOG first divides the image into blocks of \wording{50\% overlap}.
Each block contains $m\times m$ cells-per-block that each contain
$n_{pix}\times n_{pix}$ pixels-per-cell.  The computed gradient
orientation is quantized into $N_{orient}$ bins.

% Initial binning
The orientation gradient of all pixels within each cell are binned
into the quantized orientations, providing a net gradient description
within that cell.  As an example, for $N_{orient}=3$, our bins are
centered at $\theta=0, 2\pi/3, 4\pi/3$ in radians.  If a cell has no
gradient in the vertical direction, and only a gradient across the
horizontal direction, say in the positive x-direction with
$\theta=\pi/2$, it will contribute 75\% of its magnitude to the
$\theta=2\pi/3$ bin, and 25\% of its magnitude to the $\theta=0$ bin.
The bins in all cells are then concatenated to make a larger feature
vector that is $N_{orient}$\times$N_{cells}$.

% Normalization
The last step is a normalization procedure to control for illumination
effects.  Here, the sub-histograms of each cell within the same block
are normalized with respect to one another before the transformation
returns the final feature vector.  The division of the image and the
quantization of orientations provide a total of 3 parameters in HOG:
$N_{orient}$, cells-per-block, and pixels-per-cell.  We discuss how we
select parameters using cross-validation in
Section~\ref{sec:gridsearch}

%-------------------------------------------------%
\subsubsection{Optimized Pipeline Parameters with a Grid Search}\label{sec:gridsearch}
%-------------------------------------------------%

\begin{table}
\caption{Grid Search of Pipeline Parameters}
\begin{center}
\begin{tabular}{cccccc}
\hline \\ [-0.2cm]
N$_{orient}$ & Pixels/Cell & Cells/Block & C$_{reg}$ & HOG size & Score \\ [0.2cm]
\hline \\ [-0.2cm]
\multicolumn{6}{c}{(a) HST-like data} \\ [0.2cm]
\hline \\ [-0.2cm]
9 & (4, 4) & (3, 3) & 50. & &$0.76181\pm0.00491$\\ [0.2cm]
9 & (4, 4) & (1, 1) & 50. & &$0.83403\pm0.00547$\\ [0.2cm]
9 & (8, 8) & (3, 3) & 50. & &$0.85417\pm0.00851$\\ [0.2cm] 
9 & (8, 8) & (1, 1) & 50. & &$0.86389\pm0.00687$\\ [0.2cm] 
9 & (16, 16) & (1, 1) & 50. & & $0.88542\pm0.00680$ \\ [0.2cm]
9 & (16, 16) & (3, 3) & 10. & &$0.89306\pm0.00687$ \\ [0.2cm]
8 & (16, 16) & (3, 3) & 50. & &$0.90000\pm0.00170$ \\ [0.2cm]
9 & (16, 16) & (3, 3) & 50. & &$0.90139\pm0.00804$ \\ [0.2cm]
\hline \\ [-0.2cm]
\multicolumn{6}{c}{(b) LSST-like data} \\ [0.2cm]
\hline \\ [-0.2cm]
& & & & &\\
& & & & &\\ [0.2cm]
\hline
\end{tabular}
\end{center}
\label{tab:gridsearch}
\tablecomments{Panel (a) shows the results of a grid search across a
  range of HOG parameters and regression coefficient \wording{with
    training and test sets of size 1440 with the HST-like mock
    dataset.}  Panel (b) shows the corresponding results for training
  and test sets of \wording{size 5760 with the LSST mock dataset.}  We
  explore different HOG parameters in each dataset due to resolution
  and image size differences.}
\end{table}

We run a grid search across parameters that should reasonably sample
the arc edges in either the HST- or LSST-like mock observations, and
illustrate the results in Table~\ref{tab:gridsearch}.  Recall, the
HST-like images are $300\times300$ pixels per image, while the
LSST-like mock observations are $45\times45$ pixels per image.

We first estimate the size of a cell that will contain a coherent arc
feature.  To first order approximation, subdivisions of cells that are
1/100th the area of the entire image should contain coherent arc edges
that span an elongated shape within arc-containing cells.  Therefore,
\wording{we perturb the pixels per cell parameter about (30, 30) for
  the HST-like images, and (4, 4) for the LSST-like images.}

Next, the cells per block parameter determines the normalization of
each cell with respect to the neighboring cell.  In general, this will
downweight arc-like edges in cells that neighbor very bright cells,
such as cells that cover the central lens galaxy.  \wording{We therefore
  vary the cells per block parameter between (1, 1) and (3, 3).}

The number of orientations will determine the sampling of rounded
edges.  For example, if we only have two orientations, an arc-like
feature in a cell directly north of the lensing galaxy will appear in
our HOG visualization as a strong horizontal line (e.g. see top left
in Figure~\ref{fig:hogimages}), and an arc-like feature north-east of
the lensing galaxy will appear as an L-shape.  However, contributions
from a cluster or line-of-sight galaxy in the same cell will tend to
contribute edges in all orientations of the histogram (e.g. bottom
right in Figure~\ref{fig:hogimages}).

Finally, the resolution of the overall image will also limit the
additional information that an increase in $N_\text{orient}$ will
provide.  \wording{For the LSST-like data, the optimal orientation is
  $N_\text{orient}=?$ compared with $N_\text{orient}=?$ only results
  in a $???\%$ improvement.}

% Impact on image performance
The image size affects the length of the HOG feature vector, which has
a monotonically increasing relationship with the time required to
train the model.  Additionally, for fixed memory restrictions, there
is a tradeoff between the length of the feature vector and the size of
the training set.  We will discuss how the training set size affects
the train time for each data set in Section~\ref{sec:trainsetsize}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure*}[t]\label{fig:hogimages}
\begin{center}
\includegraphics[width=2\columnwidth]{figures/tiles/compilations/hog_telescope_compilation.pdf}
\caption{Left to right show example image transforms of mock images
  from Figure~\ref{fig:mockimages} with a visualized histogram of
  oriented gradients.  The image transform picks up edge features,
  with arc features showing up as edges across radial orientations.}
\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%-------------------------------------------------%
\subsection{Machine Learning Algorithm: Logistic Regression}\label{sec:LR}
%-------------------------------------------------%
\todo{ Logisitic regression (LR) is a linear classifier, which finds
  the optimal hyperplane between two classes of data points.  In our
  case, the two classes are lens and non-lens containing images.  The
  data points are the HOG feature vectors transformed from the
  training set data.  Each point occupies an N-dimensional space,
  where N is the length of the HOG vector.  In fitting for the best
  hyperplane, we must regularize the model to prevent overfitting,
  adding an additional parameter in our parameter search,
  regularization parameter $C_\text{LogReg}$.

Regularization minimizes the a cost function with respect to ....
Larger values of $C_\text{LogReg}$ correspond to increasing model
complexity.  Figure~\ref{fig:regularization} shows the accuracy of a
model for selected HOG parameters as a function of the regularization
parameter, $C_\text{LogReg}$.

The performance of a model can be measured as either the score of the
model on a test set, or some other metric, such as the area under the
Receiver Operating Characteristic curve (see Section~\ref{sec:ROC}).

The $C_{LogReg}$ will also increase the training time, \todo{since
  larger coefficients correspond to a more singular matrix}.  We
discuss the score and train time dependence on logistic regression
coefficient in Section\ref{sec:regularization}.
}
% Regularization coefficient:
% http://openclassroom.stanford.edu/MainFolder/DocumentPage.php?course=MachineLearning&doc=exercises/ex5/ex5.html

%-------------------------------------------------%
\section{Results}
\label{sec:results}
%-------------------------------------------------%

%-------------------------------------------------%
\subsection{Receiver Operating Characteristic}\label{sec:ROC}
%-------------------------------------------------%

@@

%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
%% \begin{figure}[t]\label{fig:ROCHST}
%% \begin{center}
%% \includegraphics[width=\columnwidth]{figures/ROC_HST_HOG.png}
%% \caption{Red, green, blue: ROC curves for a models trained on the
%%   HST-like data with the Logistic Regression algorithm.  The magenta
%%   solid (dotted) curve show the ROC curve for the LSST-like data with
%%   (without) the BCG.  The corresponding parameters are labeled: N for
%%   number of orientations, ppc for pixels-per-cell, cpb for
%%   cells-per-block, and C for best regularization parameter.  The AUC
%%   for each model is also labeled in the legend, where an AUC close to
%%   1 is a near-ideal model, and an AUC close to 0.5 is a model that
%%   performs as well as a coin-flip.}
%% \end{center}
%% \end{figure}
%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%            
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]\label{fig:ROCcompilation}
\begin{center}
\includegraphics[width=\columnwidth]{figures/metrics/ROC/ROC_compilation.pdf}
\caption{Red, blue, green: ROC curves for models trained on our whole
  10,000 training set and tested on our holdout set of 1,000.  These
  respectively correspond to the HST, LSST 1 year, and LSST 10 year
  data.  The solid lines are for data that include the lensing central
  galaxy, and the dashed lines for the data where there is no lensing
  galaxy, mimicking an ideal removal of the lens.  Model performance
  can be summarized by the area under the curve (AUC), labeled in the
  legend. $AUC=1$ is a perfect model, and $AUC=0.5$ is a useless
  model.}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%            
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]\label{fig:PRcompilation}
\begin{center}
\includegraphics[width=\columnwidth]{figures/metrics/PR/PR_compilation.pdf}
\caption{Red, blue, green: Precision-recall (PR) curves for models
  trained on our whole 10,000 training set and tested on our holdout
  set of 1,000.  These respectively correspond to the same models and
  data shown in Figure~\ref{fig:ROCcompilation}.  An ideal model would
  reach both a precision (purity) and recall (completeness) that equal
  1.  Note, that this performance describes a data set with a 50-50
  split between lens and non-lens containing images.}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% Logistic regression and regularization
As described in Section~\ref{sec:LR}, LR trains a model with
complexity determined by the regularization parameter coefficient,
$C_\text{LogReg}$.  Increasing model complexity (larger
$C_\text{LogReg}$), will better describe features in the training
set[wc?].  However, an overly complex model will overfit the training
set at the expense of its performance on any independent test set.
The regularization parameter ultimately defines the model performance,
and we must perform a parameter search to identify the optimal value
for $C_\text{LogReg}$.

@@

%-------------------------------------------------%
\subsection{Effects of Dataset Size on Model Performance}\label{sec:datasize}
%-------------------------------------------------% 
%-------------------------------------------------% 
\subsubsection{Effects of Training Set Size}\label{sec:trainsetsize}
%-------------------------------------------------% 
In this section, we show the effects of training set size on model
performance.  Figure~\ref{fig:trainsize} shows the AUC and the train
time depend on the log of the size of the training set.  In solid blue
lines, we see that the AUC improves almost linearly with the log of
the training size set.  In dotted red, we see that the train time
roughly scales logarithmically with the log of the train size, with an
exception of the data point corresponding to train size of 2000.

@@

%% \caption{LSST-like images with a BCG. These illustrate the same image
%%   classification paradigms as Figure~\ref{fig:HSTclassificationtile}.
%%   Note, the larger PSF of LSST reduces the resolution of images,
%%   making the arc-like morphology of faint images less significant and
%%   increases the appearance of image-BCG blending.}
%% \label{fig:LSSTBCGclassificationtile}
%% \end{figure*}

%% \begin{figure*}[t]
%%   \centering

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%            
\subsection{Performance dependence on lens-model parameters}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
%-------------------------------------------------%
\subsubsection{Effects of Image Magnification on Scoring}
\label{sec:peformance_lensmodel}
%-------------------------------------------------%
